\bf A Hierarchical Bayesian Approach for Modeling Infant-Mortality and Wearout Failure Modes

### 2017-08-23 paper/paper.tex
Replace y with t, make corrections based on feedback to Background, GLFP model

\subsection{Weibull parameterization}
\label{sec:Weibull parameterization}
For a random variable $Y$ following a Weibull distribution, the standard parameterization of the cumulative distribution function, in terms of scale parameter $\alpha$ and shape parameter $\beta$, is

$$ P(Y \le y) = 1 - \exp \left\{ \left( \frac{y}{\alpha} \right)^\beta \right\}. $$

An alternative parameterization, in terms of log-location parameter $\mu = \log \alpha$ and log-scale parameter $\sigma = 1/\beta$ is

\begin{equation} P(Y \le y) = 1 - \exp \left\{ -\exp \left( \frac{ \log y - \mu }{\sigma} \right) \right\}\\
= \Phi_{SEV}\left(\frac{\log y - \mu}{\sigma}\right),
\end{equation}
where $\Phi_{SEV}$ is the cdf of the standard smallest extreme value distribution (SEV). Thus, the Weibull distribution is a log-location scale family, with $\log Y$ following an SEV distribution with location parameter $\mu$ and scale parameter $\sigma$. With this in mind, the corresponding density function is
\begin{equation}
p(y) = \frac{1}{\sigma y}\exp\left\{\frac{\log y - \mu}{\sigma} - \exp \left( \frac{\log y - \mu}{\sigma} \right)\right\}\\
=\frac{1}{\sigma y} \phi_{SEV}\left(\frac{\log y - \mu}{\sigma}\right)
\end{equation}

An advantage conferred by writing the model in this way is that the log of the $p^{th}$ quantile, $t_p$, is a linear function of the parameters, which allows an analyst to apply prior information in a straightforward way.  Using the inverse cdf of the SEV distribution,
\begin{equation}
\log t_p = \mu + \sigma \Phi_{SEV}^{-1}(p),
\end{equation}
where $t_p = \argmin\{y:\Phi_{SEV}(\log y)>p\}$. Since $\Phi(0)\approx 0.632$, $\log{\eta} = \mu  =t_{.63}$, and when there is more prior knowledge about a quantile other than $t_{.63}$, as is often the case with high-reliability products,  this representation allows the analyst to put prior distributions directly on those quantities.

\subsection{GLFP model}
\label{subsec:GLFP model}
Following \cite{chan}, let $F_1,F_2$ be Weibull distributions with parameters $(\mu_1,\sigma_1)$ and $(\mu_2, \sigma_2)$ respectively.
The Generalized Limited Failure Population model (GLFP) of \citet{chan} is defined as follows: Let $Y \sim \op{GLFP}(\pi, \mu_1,\sigma_1,\mu_2,\sigma_2)$. Then
$$P(Y \le y) = F(y) = 1 - (1-\pi\, F_{1}(y))(1 - F_{2}(y)),\, t>0,\, 0 < \pi < 1.$$

The GLFP model can be understood as a marginalized mixture model with a binary latent variable, $\delta_i\ind \op{Bernoulli}(\pi)$. $\delta_i$ is an indicator for whether or not an observational unit is defective. Expressed conditional on $\delta_i$,
\begin{align*}
P(Y\le y | \delta_i=1) =& 1 -(1-F_1(y))(1-F_2(y))\\
P(Y\le y | \delta_i=0) =& F_2(y).
\end{align*}

The parameter $\pi_d$ represents the proportion of units susceptible to early failure, hence susceptible to both failure modes. Here the cause of failure is not assumed to be known, thus units from the same population are exchangeable.

Taking the derivative of the (marginal) cdf, the density for the GLFP is
\begin{equation}
f(y|\pi, \mu_1,\sigma_1, \mu_2, \sigma_2) = \pi f_1(y|\mu_1,\sigma_1)\left(1-F_2(y|\mu_2,\sigma_2)\right) + f_2(y|\mu_2,\sigma_2)\left(1-\pi F_2(y|\mu_2,\sigma_2\right).
\end{equation}

@@

\subsubsection{Likelihood}
We now give the general form for the likelihood function, taking into account left truncation and right censoring.  Let $Y_{i}$ denote the end of the observed lifetime of unit $i$, in hours.  Let $t_i^L$ be the left truncation time; that is, the first time reported for the unit.  Additionally, let $c_i$ be an indicator for censoring; $c_i=1$ if the failure time is right-censored, $c_1=0$ if the unit failed (at time $Y_i$). The likelihood for the GLFP is a function of $\bm{\theta} = (\mu_1, \sigma_1, \mu_2, \sigma_2, \pi)$.  Assuming the lifetimes of all units are independent, the likelihood for the data, $Y_1,\ldots,Y_N$ is given by
\begin{equation*}
L(\bm{\theta})= \prod_{i=1}^{n} \left[\frac{f(y_i;\bm{\theta})}{1-F(t_i^L;\bm{\theta})}\right]^{1-c_i} \left[ \frac{1-F(y_i;\bm{\theta})}{1-F(t_i^L;\bm{\theta})} \right]^{c_i}
\end{equation*}

\subsection{Example 1}
\label{subsec:ex1}
We fit the GLFP model on real hard drive failure data from a cloud-based storage company.  Computer hard drives are known to have a mixture of early and late failures, making the GLFP an appropriate parametric model \cite{chan}.  The data contains the number of hours the hard drive was in operation, and whether the drive failed or was right censored.  No information is available about the cause of failure.  We first focus on the drive brand model with the most observed failures to demonstrate the GLFP model.  We fit the GLFP using a Bayesian framework and specified the following proper priors to ensure a proper posterior distribution.
$$t_{0.5,1} \sim \op{N}(8,4), \quad t_{0.2,2} \sim \op{N}(10,4)$$
$$\mbox{logit}^{-1}\pi_1 \sim \op{Normal}(-3,2),\quad \sigma_1 \sim \op{Lognormal}(0, 2.5), \quad \sigma_2 \sim \op{Lognormal}(0, 2.5) $$

>>>>

\subsection{Weibull parameterization}
\label{sec:Weibull parameterization}
For a random variable $T$ following a Weibull distribution, the standard parameterization of the cumulative distribution function, in terms of scale parameter $\alpha$ and shape parameter $\beta$, is

$$ P(T \le t) = 1 - \exp \left[ \left( -\frac{t}{\alpha} \right)^\beta \right]. $$

An alternative parameterization, in terms of log-location parameter $\mu = \log (\alpha)$ and log-scale parameter $\sigma = 1/\beta$ is

\begin{equation} P(T \le t) = 1 - \exp \left\{ -\exp \left[ \frac{ \log (t) - \mu }{\sigma} \right] \right\}\\
= \Phi_{SEV}\left[\frac{\log (t) - \mu}{\sigma}\right],
\end{equation}
where $\Phi_{SEV}$ is the cdf of the standard smallest extreme value distribution (SEV). Thus, the Weibull distribution is a log-location scale family, with $\log(T)$ following an SEV distribution with location parameter $\mu$ and scale parameter $\sigma$. With this in mind, the corresponding density function is
\begin{equation}
p(t) = \frac{1}{\sigma t}\exp\left\{\frac{\log(t) - \mu}{\sigma} - \exp \left[ \frac{\log(t) - \mu}{\sigma} \right]\right\}\\
=\frac{1}{\sigma t} \phi_{SEV}\left[\frac{\log(t) - \mu}{\sigma}\right]
\end{equation}

An advantage conferred by writing the model in this way is that the log of the $p^{th}$ quantile, $t_p$, is a linear function of the parameters, which allows an analyst to apply prior information in a straightforward way.  Using the inverse cdf of the SEV distribution,
\begin{equation}
\log(t_p) = \mu + \sigma \Phi_{SEV}^{-1}(p).
\end{equation}
Because $\Phi(0)\approx 0.632$, we have $\log{\eta} = \mu  =t_{0.63}$. When there is more prior knowledge about a lower quantile, as is often the case with high-reliability products, using this parameterization allows the analyst to put prior distributions directly on those quantities. In addition, in the presence of censoring, $p$ can be chosen to make $\log(t_p)$ and $\sigma$ approximately independent. Since the choice of independent priors is common in practice, this observation lends further credence to the use of these reparameterizations.

\subsection{GLFP model}
\label{subsec:GLFP model}
Let $F_1,F_2$ be Weibull distributions with parameters $(\mu_1,\sigma_1)$ and $(\mu_2, \sigma_2)$, respectively.
The Generalized Limited Failure Population model (GLFP) of \citet{chan} is defined as follows: Let $T \sim \op{GLFP}(\pi, \mu_1,\sigma_1,\mu_2,\sigma_2)$. Then
$$P(T \le t) = F(t) = 1 - (1-\pi\, F_{1}(t))(1 - F_{2}(t)),\, t>0,\, 0 < \pi < 1.$$

The GLFP model can be understood as a marginalized mixture model with a binary latent variable, $\delta_i\ind \op{Bernoulli}(\pi)$. $\delta_i$ is an indicator for whether or not an observational unit is defective. Expressed conditional on $\delta_i$,
\begin{align*}
P(T\le t | \delta_i=1) =& 1 -(1-F_1(t))(1-F_2(t))\\
P(T\le t | \delta_i=0) =& F_2(t).
\end{align*}

The parameter $\pi$ represents the proportion of units susceptible to early failure, hence susceptible to both failure modes. Here the cause of failure is not assumed to be known, thus units from the same population are exchangeable.

Taking the derivative of the (marginal) cdf, the density for the GLFP is
\begin{equation}
f(t|\pi, \mu_1,\sigma_1, \mu_2, \sigma_2) = \pi f_1(t|\mu_1,\sigma_1)\left(1-F_2(t|\mu_2,\sigma_2)\right) + f_2(t|\mu_2,\sigma_2)\left(1-\pi F_2(t|\mu_2,\sigma_2\right).
\end{equation}

@@

\subsubsection{Likelihood}
We now give the general form for the likelihood function, taking into account left truncation and right censoring.  Let $t_{i}$ denote the end of the observed lifetime of unit $i$, in hours.  Let $t_i^L$ be the left truncation time; that is, the first time reported for the unit.  Additionally, let $c_i$ be an indicator for censoring; $c_i=1$ if the failure time is right-censored, $c_1=0$ if the unit failed (at time $Y_i$). The likelihood for the GLFP is a function of $\bm{\theta} = (\mu_1, \sigma_1, \mu_2, \sigma_2, \pi)$.  Assuming the lifetimes of all units are independent, the likelihood for the data, $Y_1,\ldots,Y_N$ is given by
\begin{equation*}
L(\bm{\theta})= \prod_{i=1}^{n} \left[\frac{f(y_i;\bm{\theta})}{1-F(t_i^L;\bm{\theta})}\right]^{1-c_i} \left[ \frac{1-F(y_i;\bm{\theta})}{1-F(t_i^L;\bm{\theta})} \right]^{c_i}
\end{equation*}

\subsection{Example 1}
\label{subsec:ex1}
We fit the GLFP model on real hard drive failure data from a cloud-based storage company.  Computer hard drives are known to have a mixture of early and late failures, making the GLFP an appropriate parametric model.  The data contain the number of hours the hard drive was in operation, and whether the drive failed or was right censored.  No information is available about the cause of failure.  We first focus on the drive-model with the most observed failures to demonstrate the GLFP model.  We the GLFP using a Bayesian computational approach, after specifying prior distributions on the model parameters. We selected proper prior distributions to improve identification of the model parameters, which ensures a proper posterior distribution. We reparameterize the component Weibull distributions using the $50^{th}$ percentile for the infant mortality failure mode ($t_{0.5,1}) and the $20^{th}$ percentile for the wearout failure mode ($t_{0.2,2}).
$$t_{0.5,1} \sim \op{N}(8,4), \quad t_{0.2,2} \sim \op{N}(10,4)$$
$$\mbox{logit}^{-1}\pi_1 \sim \op{Normal}(-3,2),\quad \sigma_1 \sim \op{Lognormal}(0, 2.5), \quad \sigma_2 \sim \op{Lognormal}(0, 2.5) $$
